---
title: "Lab 02 - Plastic waste"
author: "Otobo"
date: "1/31/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
# Load the tidyverse package
library(tidyverse)
```

```{r load-data, message=FALSE}
plastic_waste <- plastic_waste <- read.csv("C:/Users/Otobo/GitHub/progdata-class/data/data.csv")

```

## Exercises

### Exercise 1

Overall, the plastic waste per capita for countries in all continents were less than 1.0 kg per person with exception of North America where at least one country (Trinidad and Tobago) had a plastic waster per capita that great greater than 3.0kg per person.

```{r plastic-waste-continent}
#1
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  geom_histogram(binwidth = 0.2) +
  facet_wrap(~continent)

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  geom_density(binwidth = 0.2) +
  facet_wrap(~continent)


ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  geom_histogram(binwidth = 0.2) +
  geom_density(aes(y = after_stat(count)), binwidth = 0.2) +
  facet_wrap(~continent)

#2.
ggplot(plastic_waste) +
  aes(x= plastic_waste_per_cap, color = continent) + 
  geom_density()


ggplot(plastic_waste) +
  aes(x= plastic_waste_per_cap, color = continent, fill = continent, , alpha = 0.5) + 
  geom_density(alpha = 0.6)

# 3. 
# We defined color and fill of curves by mapping to the aesthetics to ensure that the boundaries for each continent have unique colors and the spaces within them are filled with the respective colors. However, we define alpha level to the geom to alter the level of opacity/transparency for each  boundary in order to make them distinguishable. Applying the alpha level to the aes() in this case doesn't result in the desired change in color.
```


### Exercise 2

```{r plastic-waste-density}

```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

```
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
```

### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements


